---
title: "How To Connect Rstudio to MYSQL To Read Data"
author: "Nnaemeka Okereafor"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About The movie rating dataset:
The code below typically demonstrate how to connect Rstudio to MYSQL database to read data from it. the dataset to read is a movie rating dataset with 44922 observations and 5 variables. To know more about the movie dataset, visit <https://www.kaggle.com/rounakbanik/the-movies-dataset>.


## Databse Connection

```{r message=FALSE, warning=FALSE}
library(RODBC)
library(tidyverse)
library("lubridate")
conn1 <- odbcConnect("MYSQL_NEWMAN")
movie_data <- sqlQuery(conn1,"SELECT * FROM movie")
head(movie_data, n=25)
```

## Dataset Dimension

```{r}
dim(movie_data)

```

## Names of Movies

```{r}

movies<-unique(movie_data$Title)
movies

```

## Adding Derived variables to original movie rating dataset

```{r}

movie_data$Review_date <- as_datetime(movie_data$Tymestamp)
movie_data$Weekday <- weekdays(movie_data$Review_date)
movie_data$Year <- year(movie_data$Review_date)
movie_data$Hour <- hour(movie_data$Review_date)
head(movie_data, n=25)

```